Synthesizing Theories of Human Language with Bayesian Program Induction

### 2018-06-20 manuscript.tex
discussion

We frame our approach as Bayesian Program Learning (BPL:
see~\cite{lake2015human}), where the model explains a set of
utterance/meaning pairs $\left\{(u_n,m_n) \right\}_{n = 1}^N$ by
inferring a theory $T$, which we model as a program.  Formalizing
grammars (theories) as generative programs has a long history in
linguistics~\cite{chomsky1968sound},
for an intuitive reason: being procedural,
programs can capture the causal nature of
both grammars and theories.
Written as a probabilistic
inference problem, the model seeks the theory $T$ maximizing
$P(\{u_n,m_n\}_{n = 1}^N | T)P(T|UG)$, where $UG$ is a ``universal
grammar'' encapsulating  higher-level abstract knowledge across
different languages.  In this BPL setting we model $UG$ as a prior
distribution over theories (programs), and represent programs as
Context-Sensitive Rewrites, a Turing-complete program representation,
but restrict the rewrites in such a way as to make them equivalent to
finite state transducers --- this approach comes from the computational
linguistics literature~\cite{beesley2003finite}.

@@

Although this framing captures the problem a BPL theory inductor needs to solve,
it offers no guidance on how to solve that problem:
the space of all programs (theories) is infinitely large and
sharply discontinuous, lacking the local smoothness that
enables local optimization algorithms (e.g., gradient descent; MCMC)
to succeed.
We adopt a strategy currently unconventional to the
AI community: constraint--based program synthesis,
where the optimization problem is translated
into a combinatorial constraint satisfaction 
problem and solved using
a SMT solver~\cite{solar2006combinatorial}.
To scale these solvers to
large and complex theories 
we wrap the solver in
an outer loop that incrementally
introduces new (utterance, meaning) pairs,
incrementally modifying the theory to explain
new data points (Supplementary materials).

@@

No theory is built from scratch: instead, researchers borrow concepts
and constraints from other successful theories.  Linguistics has long
acknowledged the importance of constraints and inductive biases in
language learning, and these principles collectively go by the name
Universal Grammar: innate for child language learners and discovered
empirically by linguists.
In practice,
both children and linguists
may have very little data to go off,
and thus success often hinges upon
having the right kind of high-level knowledge.

@@

>>>>

We frame our approach as Bayesian Program Learning (BPL:
see~\cite{lake2015human}), where the model explains a set of
utterance/meaning pairs $\left\{(u_n,m_n) \right\}_{n = 1}^N$ by
inferring a theory $T$, which we model as a program.  Formalizing
grammars (theories) as generative programs has a long history in
linguistics~\cite{chomsky1968sound},
for two intuitive reasons: being procedural,
programs can capture the causal nature of
both grammars and theories;
and being highly structured,
programs can be interpreted by human scientists.
Written as a probabilistic
inference problem, our model seeks the theory $T$ maximizing
$P(\{u_n,m_n\}_{n = 1}^N | T)P(T|UG)$, where $UG$ is a ``universal
grammar'' encapsulating  higher-level abstract knowledge across
different languages.  In this BPL setting we model $UG$ as a prior
distribution over theories (programs), and represent programs as
Context-Sensitive Rewrites, a Turing-complete program representation,
but restrict the rewrites in such a way as to make them equivalent to
finite state transducers --- this approach comes from the computational
linguistics literature~\cite{beesley2003finite}.

@@

Although this framing captures the problem a BPL theory inductor needs to solve,
it offers no guidance on how to solve that problem:
the space of all programs (theories) is infinitely large and
sharply discontinuous, lacking the local smoothness that
enables local optimization algorithms (e.g., gradient descent; MCMC)
to succeed.
We adopt a strategy based on constraint--based program synthesis,
where the optimization problem is translated
into a combinatorial constraint satisfaction 
problem and solved using
a SMT solver~\cite{solar2006combinatorial}.
To scale these solvers to
large and complex theories 
we wrap the solver in
an outer loop that incrementally
introduces new (utterance, meaning) pairs,
incrementally modifying the theory to explain
new data points (Supplementary materials).

@@

No theory is built from scratch: instead, researchers borrow concepts
and constraints from other successful theories.  Linguistics has long
acknowledged the importance of constraints and inductive biases in
language learning, and these principles collectively go by the name
Universal Grammar: innate for child language learners and 
empirically probed by linguists.
In practice,
both children and linguists
may have very little data to go off,
and thus success often hinges upon
having the right kind of high-level knowledge.

@@

Theory induction is a grand challenge for AI,
and our work here
captures only small slices of
the theory building process.
Like our model, human theorists
do craft models
by examining experimental data,
but also propose new theories by
unifying existing theoretical frameworks,
performing `thought experiments',
and inventing new formalisms.
Humans also deploy their theories more richly than our model:
proposing new experiments to test theoretical predictions,
engineering new tools based on the conclusions of a theory,
and distilling higher-level knowledge that goes far beyond what our
Fragment-Grammar approximation can represent.
Continuing to push theory induction along these many dimensions 
remains a prime target for future research.
